---
title: "unreg"
author: "Mark Wheldon"
date: "`r Sys.Date()`"
output:
    function(...) rmarkdown::html_vignette(..., css=c(system.file("rmarkdown", "templates", "html_vignette", "resources", "vignette.css", package = "rmarkdown"), "vignette2.css")):
      toc: true
bibliography: bibliography.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{unreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

 
*unreg* provides convenience functions for working with country and country classification schemes, and their ISO codes [@united_nations_standard_2019]. It is based on the package [@division19_wpp2019], specifically the data set `UNlocations` in that package.


# Motivation

Many demographic indicators are presented by country and geographic region. Regions are sets (or aggregates) of countries defined in a particular classification scheme, or 'region family'. Region families in common use are the *M49* [@united_nations_standard_2019] and *World Bank* Income Group families [@bank19_world_bank_count_lendin_group]. 

Mappings between countries and regions used by the *World Population Prospects* [@united_nations_world_2019] are given in the `UNlocations` data frame in the *wpp2019* package [@division19_wpp2019]. However, this file is quite complicated. In particular, the coding of many regions is done through specially named columns.

```{r}
library(wpp2019)
data(UNlocations)
UNlocations[1:3,1:8]
```

*unreg* provides many convenience functions that make working with the country and region names and codes much easier.


# Region Families

Region families are either hierarchical or binary. Several families of each kind are supported. 

## Hierarchical Families

Hierarchical familes have a completely nested structure. All hierarchical families have one or two levels.

M49
  : Two-level geographic grouping with two levels, e.g., "Africa", "Western Africa", ... . See [@united_nations_standard_2019].
  
SDG-Region
  : Two-level geographic grouping with two levels, e.g., "Sub-Saharan Africa", "Western Africa", ... . These were used in a 2017 report to track progress towards the *Sustainable Development Goals* of the United Nations [@united_nations_statistics_division_sdg_2019]. Following WPP 2019 [@division19_wpp2019], *unreg* further disaggregates sub-Saharang Africa in the relevant subregions from the M49 standard.
 
World Bank Income Groups
  : One-level geographic grouping [Not yet implemented]
  
Legacy Development Groups
  : [Not yet implemented, "more developed", "less developed", etc.] --- maybe might not be strictly hierarchical
  
  
## Binary Families
  
Binary families have only partially nested structures, or no nesting. They are represented by binary indicator variables which include or exclude countries.
  
SDG-Other
  : [Not yet implemented, but includes "SIDS", "LDCs", "LLDCs"]

**Have to sort through this if continue with development.**


# Package Operation

## Function Groups

Three groups of functions are provided, each for addressing a specific task:

1. Simple listing functions
2. Translation functions for translating between names and codes
3. Tabulation functions for creating easy-to-read tables of countries and regions


## Listing Functions

These functions are designed to list all available country and region names and codes. Their names all begin with `list_`.

```{r}
#### List all names---countries and regions

head(unreg::list_names())

#### List all codes---countries and regions

head(unreg::list_codes())
```


## Translation Functions

These functions translate names (countries and regions) into their ISO codes and vice-versa. They are designed to return a vector of names or codes. Their results could be looped over, for example, or for interactive use.

```{r}
#### Translate a country/region code into a name
unreg::name(c(250, 903))

#### Go the other way
unreg::code(c("France", "Africa"))

#### All names are case insensitive
unreg::code(c("france", "FRanCe"))

#### Some regions are in two families--- specify the family
unreg::code(c("Latin America and the Caribbean", "Europe", "Northern America"))
unreg::code(c("Latin America and the Caribbean", "Europe", "Northern America"),
            family = "SDG")
unreg::code(c("Latin America and the Caribbean", "Europe", "Northern America"),
            family = "M49")

#### You can abbreviate LAC
unreg:::code("lac", "SDG")
```

If you want to get the region a country belongs to use `unreg::reg_code()` or `unreg::reg_name()`. You can supply the country as its code or its name (case insensitive).

```{r}
#### Default family is "M49"
unreg::reg_code(250)
unreg::reg_code(c("France", "spAIn"))

unreg::reg_code(250, family = "SDG")

#### Default level is 1
#### Give the level to get specific level of region (can be supplied as numeric for convenience)
unreg::reg_code(250, 2, family = "SDG")

#### If you want the name, instead of the code, use 'unreg::reg_name'
unreg::reg_name("france", family = "SDG")
unreg::reg_name(250, 2, family = "SDG")
```

To get the mappings between countries and regions, use `unreg::country_codes()` and `unreg::country_names()`.

```{r}
unreg::country_codes("Africa")

unreg::country_names("Northern America")
unreg::country_names(905)       #unreg::code("Northern America", family = "M49") is 905
```


## Tabulation Functions

The tabulation functions are designed to return tables as *R* data frames that show countries and regions.

```{r} 
#### Level 1
head(unreg::reg_table(level = 1, family = "M49"))

#### Level 2
head(unreg::reg_table(level = 2, family = "M49"))


#### Both
head(unreg::reg_table(level = 1:2, family = "M49"))
#### (alternative)
head(unreg::reg_table(level = "", family = "M49"))
```


# References {-}

